---
title: "recla"
author: "Brian S. Yandell"
date: "12/8/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This has only part of the AttieDO (version 1) data. See the [setup_qtl2shiny.Rmd](https://github.com/byandell/DOread/blob/master/inst/DerivedData/setup_qtl2shiny.Rmd) and various routines in [R/DOread](https://github.com/byandell/DOread).
See [README.md](~/Documents/Research/attie_alan/DO/data/DerivedData/README.md)
in the `data/DerivedData` folder for further information.

```{r message=FALSE}
library(dplyr)
library(qtl2)
library(qtl2feather)
library(qtl2pattern)
library(stringr)
library(purrr)
```

```{r}
project_info <- data.frame(project = "AttieDO",
                           taxa = "CCmouse",
                           directory = "qtl2shinyData",
                           stringsAsFactors = FALSE)
```

```{r}
project_dir <- file.path(project_info$taxa, project_info$project)
```

```{r}
if(!dir.exists(project_dir)) {
  dir.create(project_dir)
}
```

## Query functions

Use `create_probs_query_func` and `create_mrna_query_func` from package `qtl2pattern` to query genotype probabilities.

```{r}
query_probs <-  qtl2pattern::create_probs_query_func("qtl2shinyData/CCmouse/AttieDO")
query_mrna   <- qtl2pattern::create_mrna_query_func("qtl2shinyData/CCmouse/AttieDO")
```

```{r}
saveRDS(query_probs, file.path(project_dir, "query_probs.rds"))
saveRDS(query_mrna, file.path(project_dir, "query_mrna.rds"))
```

## Genotype probabilities

These were created already (see above), but names of folders have shifted. Here is an adjustment to do once.

```{r eval=FALSE}
feather_dir <- file.path(project_dir, "genoprob")

# change feather element for allele probs
feather_aprobs <- readRDS(file.path(feather_dir, "feather_aprobs.rds"))
pr <- unclass(feather_aprobs)
pr$feather <- file.path(project_info$project, "genoprob/aprobs")
feather_aprobs <- qtl2pattern:::modify_object(feather_aprobs, pr)
saveRDS(feather_aprobs, file = file.path(feather_dir, "feather_aprobs.rds"))

# change feather element for allele pair probs
feather_probs <- readRDS(file.path(feather_dir, "feather_probs.rds"))
pr <- unclass(feather_probs)
pr$feather <- file.path(project_info$project, "genoprob/probs")
feather_probs <- qtl2pattern:::modify_object(feather_probs, pr)
saveRDS(feather_probs, file = file.path(feather_dir, "feather_probs.rds"))
```

```{r}
wd <- getwd()
setwd(datapath)
feather_aprobs <- readRDS(file.path(datapath, "AttieDO/genoprob", "feather_aprobs.rds"))
fst_aprobs <- qtl2feather::feather2fst_genoprob(feather_aprobs)
saveRDS(fst_aprobs, file = file.path(datapath, "AttieDO/genoprob", "fst_aprobs.rds"))
feather_probs <- readRDS(file.path(datapath, "AttieDO/genoprob", "feather_probs.rds"))
fst_probs <- qtl2feather::feather2fst_genoprob(feather_probs)
saveRDS(fst_probs, file = file.path(datapath, "AttieDO/genoprob", "fst_probs.rds"))
setwd(wd)
```

