---
title: "use SQLite"
author: "Brian S Yandell"
date: "September 9, 2016"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document shows how to use the SQLite to drill in to genes in a particular chromosome region and a particular phenotype. Here is the region:

```{r}
chr_id <- "3"
start_val <- 135
end_val <- 145
pheno_name <- "G8.3.AA_percent"
```


### Setup

Attach libraries.

```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(stringr)
  library(d3heatmap)
  library(ggplot2)
})
```

Connect SQL database.

```{r}
my_sql_db <- src_sqlite("islet_mrna.sqlite")
```

```{r}
my_sql_db
```

### Get ensembl information

Here we get ensembl ID, gene name, etc., from chromosome 3 between 135 and 145 Mbp.
Note that the `chr` must be quoted (as a character string) rather than a number.
Some remarks:

* The `()`s aroudn the assignment will print out the first few lines; not needed if you just want to save it.
* The `%>%` are pipes to limit the data from the full 21,771 entries to the 60 in this interval.
* The `collect` actually gets the data from the database so you can use it.

```{r}
(ensembl_info_3 <- tbl(my_sql_db, "ensembl_info") %>%
  filter(chr == chr_id, start >= start_val, end <= end_val) %>%
  collect)
```

### Get correlations in this interval.

Use the collected interval information to get a set of correlations.
Remember that you have to `collect` to actually get data for use.

```{r}
(islet_cors_3 <- tbl(my_sql_db, "islet_pheno_cors") %>%
  filter(ensembl_id %in% ensembl_info_3$ensembl_id) %>%
  collect)
```

#### D3 heatmap

First change into data frame with `symbol.nnn` as row name, where `symbol` is the gene name (blank if not provided) and `nnn` is the numeric part of the ENSEMBL ID.

```{r}
islet_cors_3 <- data.frame(islet_cors_3)
m <- match(islet_cors_3$ensembl_id, ensembl_info_3$ensembl_id)
row.names(islet_cors_3) <- paste(ensembl_info_3$symbol[m],
                                 str_replace(ensembl_info_3$ensembl_id,
                                             "ENSMUSG0+", ""),
                                 sep = ".")
```

```{r}
d3heatmap(islet_cors_3[,-1])
```

### Get expression and phenotype data.

#### Get just the `r pheno_name` phenotype.

```{r}
(pheno <- tbl(my_sql_db, "pheno") %>%
  select_("mouse","sex","wave",pheno_name) %>%
  collect)
```

Pick `ensemble_id`s extremes of corelation with `r pheno_name`.

```{r}
wh <- c(which.min(islet_cors_3[,pheno_name]),
        which.max(islet_cors_3[,pheno_name]))
(wh <- islet_cors_3[wh,"ensembl_id"])
```

Get expression data in this interval for these `ensembl_id`s.

```{r}
(islet_mrna_3 <- tbl(my_sql_db, "islet_mrna") %>%
  filter(ensembl_id %in% wh) %>%
  collect)
```

#### Plot phenotype against expression

First join the data

```{r}
data <- full_join(pheno, islet_mrna_3) %>%
  filter(!is.na(sex)) %>%
  mutate(wave = factor(wave))
```

Since `pheno_name` is a variable, it is a bit tricky to include it in a plot. Here we create log10 of pheno as a new column

```{r}
data$log10_pheno <- log10(data[[pheno_name]])
```

Now plot data using `symbol` and `ensembl_id` in title.

```{r}
for(ens in wh) {
  symbol <- (ensembl_info_3 %>%
               filter(ensembl_id == ens))$symbol
  print(ggplot(data %>% 
                 filter(ensembl_id == ens),
               aes(expr,log10_pheno,col=wave)) +
          geom_point() +
          geom_smooth(method = "lm", se = FALSE) +
          facet_wrap(~sex) +
          ylab(paste("log10 of pheno_name")) +
          ggtitle(paste(symbol, ens)))
}
```

